Impact ionization processes in a photodriven Mott insulator: influence of phononic dissipation

### 2022-10-31 resub_PRB/letters/reply_to_referees.txt

Dear Dr. Kancharla,

Thank you very much for your e-mail from October 24th with the reply of the Referees on our paper. The first Referee recommends publication without hesitation. The second Referee believes that the paper has to be published in Physical Review B, discussing the points he raised.

In the new version of our paper we have taken into account in detail all comments. A detailed reply is also included to this letter.

We hope that our paper is now suitable for publication in Phys. Rev. B.

Yours sincerely,
Paolo Gazzaneo, Tommaso Maria Mazzocchi, Jan Lotze and Enrico Arrigoni

--------------------------------------------------
          First Referee                          
--------------------------------------------------

1. As the authors pointed out the electric field amplitude considered
in this paper is several orders of magnitude larger than that of the
sunlight, therefore, strictly speaking the results presented in the
present paper are more applicable to intense laser-light
photoexcitations

2. The authors have made an effort to extrapolate in the regime of
smaller electric field amplitude and smaller hybridization strength.
They note that in order to achieve the latter one would need to
several layers where photexcited carriers separated by an electric
field gradient would have more time to induce impact ionization before
escaping into the leads.

3. The authors plan in future investigation to address the previously
mentioned issue.

A1-A2-A3: We thank a lot the Referee for his remarks and acknowledgements about our article.

4. I would like to suggest that, perhaps, the authors, in their future
investigations, could include the role of “bad” metallic contacts,
maybe represented by narrow-band leads or possibly with a small-width
insulating barrier between the leads and the material as abstracted by
the Hubbard-like model studied. If the insulating barrier is
microscopically small the carriers would still be able to tunnel, but
perhaps would remain longer in the region of the material which might
allow II to occur.

A4: In our future investigation, we will keep this suggestion in mind.

----------------------------------------------------
        Second Referee                              
----------------------------------------------------

************************************************************************
*** To highlight the changes in the main text we marked them in BLUE ***
************************************************************************

1. It is really important to discuss why DMFT approximation is
expected to be meaningful in quasi-2D case (1-layer Hubbard model
coupled to leads).

A1: The Referee is right, we acknowledge that a discussion on the meaninfulness of DMFT approximation for quasi-2D cases is missing.

In the quasi-2D case we are addressing, the use of DMFT and therefore neglecting the momentum dependence of the electron self-energy leads to meaningful results for two main reasons. 
First: The considered system is not "strictly" two dimensional, meaning that there is still some dispersion in the third dimension, even though weak. This is exactly the case of our setup.
Second: We are not including in our treatment long-range magnetic fluctuations or, more generally, spin-fluctuations. The inclusion of such physics gives truly nonlocal effects and therefore nonlocal contributions in the self-energy, whose momentum-dependence has to be clearly taken into account in this case. In lower dimensional systems, as the one we are considering, these effects are usually not negligible, but this depends on the physics one is interested in: to clearly address the influence of the leads and acoustic phonons on impact ionization processes, spatially nonlocal correlations would complicate the overall picture and understanding.
Regarding this point, we are also limited in considering only single orbitals in the current version of our nonequilibrium many-body impurity solver.

As asked, we added in Sec.I a discussion on this point, with a reference to a comprehensive review of nonlocal correlations beyond DMFT (Rev. Mod. Phys. 90, 025003
(2018)).

2. It seems that chosen temperature is pretty high. Please discuss why
optical phonons are neglected.

A2:

We thank the Referee for raising this point, giving us the opportunity to clarify it better.

The value of the temperature of T=0.02, is not high in comparison with the other characteristic energy scales of the system, as e.g. the soft cutoff acoustic phonon frequency and the Hubbard interaction. It is actually the smallest chosen energy scale. Usually, the optical phonons modes energy is at least twice the acoustic phonons one, therefore the choice of a small value for the temperature justifies in our view taking into account acoustic phonons and neglecting optical phonons.
Just as a remark, we are interested in the effects of phononic dissipation and, to address such issue, the acoustic phonons with a finite bandwidth, in contrast to optical phonons, are more effective in energy dissipation, as already been written in Sec.II.

As the Referee suggests, we inserted a footnote in Sec.IV, addressing explicitly this point.

3. It is naturally expected that phonons reduce the bandgap (here,
pseudogap), as in standard semiconductors, where the phonons give
10-100meV reduction of the gap and the effect increases with
temperature. Strictly speaking, to obtain the change of the
bandstructure due to phonons one needs to calculate the Fan and DW
terms of the , generally speaking, momentum-dependent electron
self-energy. In this sense, again, how important are momentum
dependencies of electron Sigma in 2D case?

A3: We thank the Referee to have pointed out a similar effect of reduction of the pseudogap in semiconductors and for the clarification about the phonon renormalization of the band structure due to momentum-dependent self energy. This is an important point and we address the issue below.

The importance of momentum dependencies of the electron self-energy in the 2D case depends on the level of sophistication one would like to achieve in considering acoustic phonons and therefore if one is more interested in quantitative results or not. As explicitly stated in Sec.I, "...our analysis can only provide qualitative results, while quantitative comparisons with experimental systems require a more realistic setup and are beyond the scope of this work."
At this level, having used DMFT approximation and therefore considered momentum-independent self-energies, we can't quantify the importance of momentum dependencies in the electron self-energy.
However, as pointed out in the review Rev. Mod. Phys. 89, 015003, the Fan term is also "...traditionally referred to as the self-energy in the Migdal approximation". This is exactly the term we included, without momentum dependence, in the Dyson equation to take into account the effect of the acoustic phonons on electrons. In the same review, it's stated that "The Debye-Waller contribution to the electron self-energy is almost invariably ignored in the literature on metals and superconductors...". In a wider view, systems displaying strong correlations, as we one taken into account, fall into this category, therefore we are confident that such term can be safely neglected for a qualitative description of the effects of acoustic phonons.

We included a footnote in Sec.IIIB addressing such issue, together with a reference to the review mentioned above (Rev. Mod. Phys. 89, 015003).

4. It is not clear from Conclusions (at least for me) what stopped the
authors to consider more realistic weaker fields and hybridizations.

A4:

We thank the Referee for having brought up this important point.

We didn't consider weaker electric field amplitudes since DMFT only converges slowly for such a choice of parameters. We have already discussed this point in the footnote [47]. The same reasoning applies for smaller hybridizations, slowing down in a considerable way the convergence of the DMFT cycle.

We followed the suggestion of the Referee, putting in the main text in Sec.V the footnote [47], extending it to mention smaller hybridizations.

List of changes:

**********************************************************************************************
*** To highlight the changes we marked them in BLUE (in source code denoted by '\resub{}') ***
**********************************************************************************************

Motivated by questions/observations from Second Referee:

Modified Sec.I  
Added reference in Sec.I
Added footnote in Sec.IIIB
Added reference in Sec.IIIB
Added footnote in Sec.IV
Modified Sec.V

### 2022-10-31 resub_PRB/Paper1.tex
added resub_PRB folder

@@

For this goal we consider the simplified model shown in Fig.~\ref{fig:setup}, consisting of a Hubbard layer located between two leads. The leads have themselves a wide band, as expected for good conductors, which is effectively narrowed due to the weak coupling to intermediate layers which we take as simplified models for the contacts. Electrons on the Hubbard layer interact locally with acoustic phonons. An external periodic electric field induces a Floquet steady state with a current flowing from the lead with the lower to the lead with the higher chemical potential, thus transferring its energy.
 We address this periodic problem via nonequilibrium Floquet dynamical mean-field theory (F-DMFT).

@@

As in Ref.~\cite{ma.ga.22u}, phonons are included in the form of acoustic  branches attached to each lattice site and coupled to the electrons via the Hamiltonian 
\begin{equation}\label{eq:e-ph_ham} 
\hat{H}_{\text{e-ph}} = g \sum_{i\sigma} \hat{n}_{i\sigma} \hat{x}_{i},  
\end{equation}
with $g$ as e-ph coupling and $\hat{x}_{i} \equiv \frac{1}{\sqrt{2}} \left( \hat{b}^{\dagger}_{i} + \hat{b}_i \right)$, where $\hat{b}^{\dagger}_{i}$ ($\hat{b}_{i}$) creates (annihilates) a phonon of the acoustic branch at site $i$, with dispersion relation described by $\hat{H}_{\text{ph}}$, discussed in Sec~\ref{sec:Dyson_equation}. In contrast to Holstein phonons, they have a finite bandwidth, which is more effective in energy dissipation. Details on the leads are provided in Sec.~\ref{sec:Dyson_equation}.

@@

The electron SE $\underline{\bf{\Sigma}}$ is obtained from F-DMFT and therefore, in this approximation, independent of the crystal momentum, i.e. $\underline{\bf{\Sigma}}(\omega;\epsilon,\overline{\epsilon}) \simeq \underline{\bf{\Sigma}}(\omega)$. Further details are given in Sec.~\ref{sec:FDMFT_implementation}. In the same spirit, the e-ph SE is included locally as $\underline{\bf{\Sigma}}_{\text{e-ph}}(\omega;\epsilon,\overline{\epsilon}) \simeq \underline{\bf{\Sigma}}_{\text{e-ph}}(\omega)$. In terms of the Keldysh contour time arguments $z$ and $z^\prime$, it has the form

@@

The Keldysh components of the non-interacting phonon GF $\kel{D}_{\text{ph}}(t,t^{\prime})$ are given by~\cite{ao.ts.14}
%
\begin{align}\label{eq:Ph_Prop_time}
\begin{split} 
D^{\text{R}}_{\text{ph}}(t,t^{\prime}) & = -\ii \theta(t-t^{\prime}) \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega), \\
D^{>}_{\text{ph}}(t,t^{\prime}) & = -\ii \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega) \left[ 1 + b(\omega) \right] \\
D^{<}_{\text{ph}}(t,t^{\prime}) & = -\ii \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega) \ b(\omega), \\
\end{split}, 
\end{align}
%
where $b(\omega)=1/(\ee^{\beta\omega}-1)$ is the Bose-Einstein distribution function at inverse temperature $\beta$. We consider acoustic phonons, with spectral function $A_{\text{ph}}(\omega) = (\omega/\omega^{2}_{\text{ph}}) \ee^{-| \omega|/\omega_{\text{ph}}}$,  $\omega_{\text{ph}}$ being a soft cutoff frequency~\cite{pi.li.21}. The retarded and Keldysh components of the e-ph SE are easily extracted from  Eq.~\eqref{eq:backbone_e-ph_SE} and can be found in Ref.~\cite{ma.ga.22u}.

@@

In order to study II in a Mott insulating layer, we adjust the bands to achieve the energy scheme shown in Fig.~\ref{fig:energy_setup}(a). 
 We follow Ref.~\cite{so.do.18} and take $U=12$ and unless stated otherwise $E_0=2$. The fermion leads and the acoustic phonons have temperature $T=0.02$. We consider a particle-hole symmetric system with $\Gamma_{\text{l}}=\Gamma_{\text{r}}=0.37$, $t_{\text{l}}=t_{\text{r}}=1.7$, $\varepsilon_{\text{l}/\text{r}}=\mp 6$ and $v_{\text{l}}=v_{\text{r}}=v$. The parameters $t_{\text{l/r}}$, $\varepsilon_{\text{l/r}}$, and  $\Gamma_{\text{l/r}}$ are chosen such that the leads' DOS approximately overlaps with the Hubbard bands, i.e. they exhibit a \textquoteleft bandwidth\textquoteright~\footnote{Notice that the support of the leads' spectra are formally infinite. $W_{\text{b}}$ is the value at which the spectrum gets  suppressed.} $W_{\text{b}}\approx 8$ and are centered at the same position. The chemical potentials are set to $\mu_{\text{l}/\text{r}} =\mp 1$, so that a current from left to right is produced by taking energy from the driving.

@@

This behavior is qualitatively valid for different values of the soft cutoff phonon frequency
 $\omega_{\text{ph}}$. It can be seen in Fig.~\ref{fig:j_vs_omega_mu1_v_0.4_0.8_E0_2_sweep_omegaph}, which shows that the impact of acoustic phonons on 
current and double occupation is slightly boosted when decreasing $\omega_{\text{ph}}$~\footnote{Reducing $\omega_{\text{ph}}$ has two effects. First, the phonon spectral function $A_{\text{ph}}(\omega)$ exhibits more weight at low frequencies around $\omega\approx\omega_{\text{ph}}$. Second, it restricts the maximum value of the reciprocal lattice vector $\vec{q}_{\text{max}}$ via $\omega_{\text{ph}}=\omega(\vec{q}_{\text{max}})$~\cite{ma.ga.22u}, meaning that short-wavelength phonons are suppressed.}.

@@

Experimentally, the electric field amplitude considered in this paper~\footnote{We don't consider smaller electric field amplitudes $E_0$ since DMFT only converges slowly for such a choice of parameters.} is several orders of magnitude larger than that of the sunlight, therefore our results are more relevant for photoexcitations produced by intense laser pulses (see also the discussion in~\cite{mu.we.18}) rather than for true photovoltaic systems. On the other hand, one may try to qualitatively extrapolate the discussion at the end of Sec.~\ref{sec:E0_2_only_electrons} in the other direction to smaller electric field amplitudes, for which a smaller hybridization would be necessary in order to sustain II. In a more realistic setup modelling oxide heterostructures~\cite{as.bl.13,pe.be.19}, a regime of effectively small hybridization would be achieved by considering several layers in which photoexcited carriers, separated by an electric field gradient, have more time to induce II processes before escaping into the leads. To address the occurrence of II in such a setup, further extensions such as considering multiple orbitals and impurity scattering should be taken into account. These would be interesting for future studies.

@@

\bibliographystyle{prsty}   % references in Phys Rev style
\bibliography{references_database}

>>>>

\definecolor{hblue}{RGB}{0,0,255}
\newcommand{\resub}[1]{{\color{hblue} #1}}

@@

For this goal we consider the simplified model shown in Fig.~\ref{fig:setup}, consisting of a Hubbard layer located between two leads. The leads have themselves a wide band, as expected for good conductors, which is effectively narrowed due to the weak coupling to intermediate layers which we take as simplified models for the contacts. Electrons on the Hubbard layer interact locally with acoustic phonons. An external periodic electric field induces a Floquet steady state with a current flowing from the lead with the lower to the lead with the higher chemical potential, thus transferring its energy.
 We address this periodic problem via nonequilibrium Floquet dynamical mean-field theory (F-DMFT) \resub{, which has been proven to give meaningful results in similar setups (see, e.g.~\cite{so.do.18}) and when spatially non-local correlations as long-range magnetic fluctuations are explicitly not taken into account~\cite{ro.ha.18}.}

@@

As in Ref.~\cite{ma.ga.22}, phonons are included in the form of acoustic  branches attached to each lattice site and coupled to the electrons via the Hamiltonian 
\begin{equation}\label{eq:e-ph_ham} 
\hat{H}_{\text{e-ph}} = g \sum_{i\sigma} \hat{n}_{i\sigma} \hat{x}_{i},  
\end{equation}
with $g$ as e-ph coupling and $\hat{x}_{i} \equiv \frac{1}{\sqrt{2}} \left( \hat{b}^{\dagger}_{i} + \hat{b}_i \right)$, where $\hat{b}^{\dagger}_{i}$ ($\hat{b}_{i}$) creates (annihilates) a phonon of the acoustic branch at site $i$, with dispersion relation described by $\hat{H}_{\text{ph}}$, discussed in Sec~\ref{sec:Dyson_equation}. In contrast to Holstein phonons, they have a finite bandwidth, which is more effective in energy dissipation. Details on the leads are provided in Sec.~\ref{sec:Dyson_equation}.

@@

The electron SE $\underline{\bf{\Sigma}}$ is obtained from F-DMFT and therefore, in this approximation, independent of the crystal momentum, i.e. $\underline{\bf{\Sigma}}(\omega;\epsilon,\overline{\epsilon}) \simeq \underline{\bf{\Sigma}}(\omega)$. Further details are given in Sec.~\ref{sec:FDMFT_implementation}. In the same spirit, the e-ph SE is included locally as $\underline{\bf{\Sigma}}_{\text{e-ph}}(\omega;\epsilon,\overline{\epsilon}) \simeq \underline{\bf{\Sigma}}_{\text{e-ph}}(\omega)$ ~\footnote{\resub{As explicitly stated in Sec.~\ref{sec:intro}, our analysis provides only qualitative results. A more accurate treatment that aims to provide quantitative predictions, should take into account the crystal momentum dependence of $\underline{\bf{\Sigma}}_{\text{e-ph}}$. We refer to the review~\cite{gius.17} for a detailed and comprehensive explanation.}}. In terms of the Keldysh contour time arguments $z$ and $z^\prime$, it has the form

@@

The Keldysh components of the non-interacting phonon GF $\kel{D}_{\text{ph}}(t,t^{\prime})$ are given by~\cite{ao.ts.14}
%
\begin{align}\label{eq:Ph_Prop_time}
\begin{split} 
D^{\text{R}}_{\text{ph}}(t,t^{\prime}) & = -\ii \theta(t-t^{\prime}) \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega), \\
D^{>}_{\text{ph}}(t,t^{\prime}) & = -\ii \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega) \left[ 1 + b(\omega) \right] \\
D^{<}_{\text{ph}}(t,t^{\prime}) & = -\ii \int \dd\omega \ \ee^{-\ii\omega\left(t-t^{\prime}\right)} A_{\text{ph}}(\omega) \ b(\omega), \\
\end{split}, 
\end{align}
%
where $b(\omega)=1/(\ee^{\beta\omega}-1)$ is the Bose-Einstein distribution function at inverse temperature $\beta$. We consider acoustic phonons, with spectral function $A_{\text{ph}}(\omega) = (\omega/\omega^{2}_{\text{ph}}) \ee^{-| \omega|/\omega_{\text{ph}}}$,  $\omega_{\text{ph}}$ being a soft cutoff frequency~\cite{pi.li.21}. The retarded and Keldysh components of the e-ph SE are easily extracted from  Eq.~\eqref{eq:backbone_e-ph_SE} and can be found in Ref.~\cite{ma.ga.22}.

@@

In order to study II in a Mott insulating layer, we adjust the bands to achieve the energy scheme shown in Fig.~\ref{fig:energy_setup}(a). 
 We follow Ref.~\cite{so.do.18} and take $U=12$ and unless stated otherwise $E_0=2$. The fermion leads and the acoustic phonons have temperature $T=0.02$~\footnote{\resub{The chosen value of the temperature is, in comparison with the other paramaters of the system, as e.g. $\omega_{\text{ph}}$ and $U$, the smallest energy scale. For this reason, we can neglect optical phonons in our qualitative treatment.}}. We consider a particle-hole symmetric system with $\Gamma_{\text{l}}=\Gamma_{\text{r}}=0.37$, $t_{\text{l}}=t_{\text{r}}=1.7$, $\varepsilon_{\text{l}/\text{r}}=\mp 6$ and $v_{\text{l}}=v_{\text{r}}=v$. The parameters $t_{\text{l/r}}$, $\varepsilon_{\text{l/r}}$, and  $\Gamma_{\text{l/r}}$ are chosen such that the leads' DOS approximately overlaps with the Hubbard bands, i.e. they exhibit a \textquoteleft bandwidth\textquoteright~\footnote{Notice that the support of the leads' spectra are formally infinite. $W_{\text{b}}$ is the value at which the spectrum gets  suppressed.} $W_{\text{b}}\approx 8$ and are centered at the same position. The chemical potentials are set to $\mu_{\text{l}/\text{r}} =\mp 1$, so that a current from left to right is produced by taking energy from the driving.

@@

This behavior is qualitatively valid for different values of the soft cutoff phonon frequency
 $\omega_{\text{ph}}$. It can be seen in Fig.~\ref{fig:j_vs_omega_mu1_v_0.4_0.8_E0_2_sweep_omegaph}, which shows that the impact of acoustic phonons on 
current and double occupation is slightly boosted when decreasing $\omega_{\text{ph}}$~\footnote{Reducing $\omega_{\text{ph}}$ has two effects. First, the phonon spectral function $A_{\text{ph}}(\omega)$ exhibits more weight at low frequencies around $\omega\approx\omega_{\text{ph}}$. Second, it restricts the maximum value of the reciprocal lattice vector $\vec{q}_{\text{max}}$ via $\omega_{\text{ph}}=\omega(\vec{q}_{\text{max}})$~\cite{ma.ga.22}, meaning that short-wavelength phonons are suppressed.}.

@@

Experimentally, the electric field amplitude considered in this paper is several orders of magnitude larger than that of the sunlight, therefore our results are more relevant for photoexcitations produced by intense laser pulses (see also the discussion in~\cite{mu.we.18}) rather than for true photovoltaic systems. \resub{We didn't consider smaller electric field amplitudes, because of the slow F-DMFT convergence for such a choice of parameters, that occurs also with weaker hybridization strengths.}

On the other hand, one may try to qualitatively extrapolate the discussion at the end of Sec.~\ref{sec:E0_2_only_electrons} in the other direction to smaller electric field amplitudes, for which a smaller hybridization would be necessary in order to sustain II. In a more realistic setup modelling oxide heterostructures~\cite{as.bl.13,pe.be.19}, a regime of effectively small hybridization would be achieved by considering several layers in which photoexcited carriers, separated by an electric field gradient, have more time to induce II processes before escaping into the leads. To address the occurrence of II in such a setup, further extensions such as considering multiple orbitals and impurity scattering should be taken into account. These would be interesting for future studies.

@@

\bibliographystyle{prsty}   % references in Phys Rev style
\bibliography{footnotes,references_database}
